Bayesian exponential random graph modeling of KWL Burial network

### 2020-10-20 analysis/paper/paper.Rmd
adjusting settings for different prior

<!--burial study- network analysis-->
Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole. With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017]. Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007]. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

@@

We analyzed burial data collected from published excavation reports, and original fieldwork notes from Kiwulan. We do not include the burials of the Chinese phase in our network analysis due to small sample size. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. In our case, burials are nodes that are linked when they have goods in common. Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

<!---Based on our hypothesis, we model a network using parameters including reciprocity, transitivity, popularity, and homophily structure effects. We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence. This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality. We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall. This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any. --->

@@

<!--Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017] -->

@@

θ1 = number of ties
θ2 = individuals with the same abundance of burial goods
θ3 = gwesp, a positive sign suggests a tendency of burials in our sample to group together in closed transitive structures. In mortuary context, clustering implies a tendency of burials with the same status indicated by prestige goods to be organised in closed structures. Burials with multiple prestige goods in common are more likely to be directed connected.
θ4 = gwdegree, negative estimates reflect an increased likelihood on ties to higher-degree nodes, or the strong edges are not necessarily centralised or dispersed in the degree distribution.

>>>>

<!--burial study- network analysis-->
Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole. With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals, allowing the inference for social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017]. Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007]. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

@@

We analyzed burial data from published excavation reports, and original fieldwork notes of Kiwulan. We do not include the burials of the Chinese phase in our network analysis due to small sample size. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. In our case, burials are nodes that are linked when they have goods in common. Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

@@

@@

θ1 = number of ties
θ2 = individuals with the same abundance of burial goods
θ3 = gwesp, a positive sign suggests a tendency of burials in our sample to group together in closed transitive structures. In mortuary context, clustering implies a tendency of burials with the same status indicated by prestige goods to be organized in closed structures. Burials with multiple prestige goods in common are more likely to be directed connected.
θ4 = gwdegree, negative estimates reflect an increased likelihood on ties to higher-degree nodes, or the strong edges are not necessarily centralised or dispersed in the degree distribution.
